Security Properties for Stack Safety

### 2021-05-14 short/main.tex
Polish the rest

We propose such a characterization
using the tools of language-based security. The informal claim that
``stack safety protects a caller from its callee'' amounts to saying
that it guarantees
the {\em integrity} and {\em confidentiality} of the caller's local state
until it regains control.
%
We formalize these integrity and confidentiality requirements as trace
(hyper-)properties, with two different
variants: {\em stepwise} variants, in which a caller's data must {\em never} be
read or modified during a call, and {\em observational} ones, in which
callees may read from and write to their caller's stack frame, as
long as these ``risky'' behaviors do not affect the system's observable
behavior. The observational properties are more extensional and thus
allow ``lazier'' protection mechanism that permit risky reads or
writes as long as they do not affect the program's visible outcomes.

Confidentiality is especially interesting.  It is based on a traditional
notion of noninterference, but whereas ordinary noninterference is 
an end-to-end hyper-property on whole program runs, stack confidentiality is
a ``nested'' form of 
noninterference applied to
subtraces of the whole program trace delimited by call instructions and
corresponding returns, requiring that the callee's behavior is invariant
under hypothetical changes to the caller's stack frame.
%behavior of the callee is
%independent of the caller's stack frame.
%, between each call
%instruction and the corresponding return, the
%caller's stack frame is protected from the callee.

This formulation not only captures the intuition
that the callee cannot directly access the caller's state, but gives a novel
way of looking at control-flow attacks. We are not interested in attacks that
merely aim to execute arbitrary code -- in fact, our attacker model assumes
that the (attacking) callee may already execute arbitrary code.
%
By extension, we do not place any particular restrictions on control flow.
Perhaps suprisingly, we can still meaningfully distinguish the caller from
its callee, and talk about protecting the caller's data. After a call, control
is considered to belong to the callee until the trace reaches a valid
``return target,'' typically a state with the program counter at the next instruction
and the stack pointer restored. If the caller's data are accessed before then,
it is a violation of integrity or confidentiality.

These properties can optionally be extended with a notion of {\em well-bracketed control flow}
as in \citet{SkorstengaardSTK}, the system-wide control-flow property that callees always return
to their immediate caller, if they return at all. This property is largely orthogonal to our
data-protection properties.

To demonstrate the utility of our properties, we use them
to evaluate an existing mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V specification. We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that Roessler and Dehon's micro-policies correctly abort programs that
would violate stack safety.
Furthermore, we check incorrect
enforcement variants (both variants that we accidentally created
during our re-implementation of the micro-policy and ones that we
intentionally crafted to be broken in order to increase our confidence
in testing and the enforcement mechanism itself). The testing framework
is able to generate counterexamples that violate our properties but are not halted by
incorrect enforcement mechanisms, increasing our confidence when it finds
no counterexamples under the real enforcement mechanisms.

Our testing supports the stack-safety claim of Roessler and Dehon's {\em Depth
  Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the depth of
the function activation that owns the frame and access to those locations is
then permitted only when that activation is currently executing.
On the other hand, we find that their \emph{Lazy Tagging and Clearing} policy
violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak happens to use the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that should enforce observational
integrity and confidentiality, albeit at some performance cost. The next step
is to expand testing to these properties and validate the variant; efficiently
testing observational properties presents unique challenges because a property
violation can be arbitrarily far removed from the initial ``risky'' read or write.

We have demonstrated our model's flexibility by extending it to different settings.
First, we can support passing arguments on the stack, allowing some of the state of
the caller to be shared with the callee (and potentially with a further
callee, and so on.)  We can also describe a simple coroutine model, with one
statically bounded stack per coroutine, each protected from the others.

In ongoing work, we are investigating whether stack protection schemes
based on the Cheri capability system~\citep{Woodruff+14,Chisnall+15}
can enforce our properties.  In particular, we are adapting 
the uninitialized capability model of \citet{Georges+21} to work with
our testing framework for validation against our properties. Of the several Cheri-based
stack protection schemes, its treatment of unitialized frames seems most compatible
with our model of confidentiality, without the need for expensive stack clearing
between calls. We expect that Cheri-based techniques support our properties if we
assume extra cooperation from callers, namely that they do not leak their capabilities
to the attacker. %There is no equivalent dynamic requirement in a tag-based system because
%the policy can dynamically enforce cooperation.

>>>>

We propose such a characterization
using the tools of language-based security. The informal claim that
``stack safety protects a caller from its callee'' amounts to saying
that it guarantees
the {\em integrity} and {\em confidentiality} of the caller's local state
until it regains control.
%
We formalize these integrity and confidentiality requirements as trace
(hyper-)properties, with two different
variants: {\em stepwise} variants, in which a caller's data must {\em never} be
read or modified during a call, and {\em observational} ones, in which
callees may read from and write to their caller's stack frame, as
long as these ``risky'' behaviors do not affect the system's observable
behavior. The observational properties are more extensional and thus
allow ``lazier'' protection mechanisms that permit risky reads or
writes as long as they do not affect visible outcomes.

Confidentiality is especially interesting.  It is based on a traditional
notion of noninterference, but whereas ordinary noninterference is
an end-to-end hyper-property on whole program runs, stack confidentiality is
a ``nested'' form of
noninterference applied to
subtraces of the whole program trace delimited by call instructions and
corresponding returns, requiring that the callee's behavior is invariant
under hypothetical changes to\bcp{scrambling of?} the caller's stack frame.
%behavior of the callee is
%independent of the caller's stack frame.
%, between each call
%instruction and the corresponding return, the
%caller's stack frame is protected from the callee.

This formulation not only captures the intuition
that the callee cannot directly access the caller's state; it also gives a novel
way of looking at control-flow attacks. We are not interested in\bcp{I find
  ``We are not interested in'' confusing here---the previous sentence
  promises a novel way of looking, and this opening feels irrelevant to
  that.} attacks that
merely aim to execute arbitrary code -- indeed, our model assumes
that attackers (i.e., callees) may already execute arbitrary code.
%
By extension, we do not place any particular restrictions on control flow.
Perhaps suprisingly, we can still meaningfully distinguish the caller from
its callee, and talk about protecting the caller's data. After a call, control
is considered to belong to the callee until the trace reaches a valid
``return target,'' typically a state with the program counter at the next instruction
and the stack pointer restored. If the caller's data are accessed before then,
it is a violation of integrity or confidentiality.  \bcp{Actually, the whole
paragraph doesn't make very much sense to me.}

These properties can optionally be extended with a notion of {\em well-bracketed control flow}
as in \citet{SkorstengaardSTK}---a global requirement that callees
must always return
to their immediate caller, if they return at all. This property is
largely\bcp{``largely'' feels like a weasel word...} orthogonal to our
data-protection properties.

To demonstrate the utility of our properties, we use them
to evaluate an existing mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V ISA model\bcp{instead of
  ``spec''.  OK?}. We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that Roessler and Dehon's micro-policies correctly abort programs that
would violate stack safety.
Our testing regime is also able to detect incorrect variants of the
enforcement mechanisms---both variants that we accidentally created
during our re-implementation of the micro-policy and ones that we
intentionally crafted to be broken in order to increase our confidence
in testing and the enforcement mechanism itself. For these broken
mechanisms, our tester
generates counterexamples that violate our properties but escape the
enforcement.  This increases our confidence in enforcement
mechanisms for which the tester is unable to find  counterexamples.

Our testing supports the stack-safety claim of Roessler and Dehon's {\em Depth
  Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the depth of
the function activation that owns the frame and access to those locations is
then permitted only when that activation is currently executing.
On the other hand, we find that their \emph{Lazy Tagging and Clearing} policy
violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak happens to use the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that should\bcp{``that
  we believe will''?} enforce observational
integrity and confidentiality, albeit at some performance cost. (An obvious next step
is to expand testing to these properties and validate the variant, but
efficiently
testing observational properties is more challenging because a property
violation can be arbitrarily far removed from the initial ``risky'' read or
write.)

We have demonstrated\bcp{let's use the same verb tense consistently---''we
  do X'' [my preference] or ``we have done X'' or ``we did X''---all the way
through} our model's flexibility by extending it to different settings.
First, we can support passing arguments on the stack, allowing some of the state of
the caller to be shared with the callee (and potentially with a sub-callee,
and so on.)  We can also describe a simple coroutine model, with one
statically bounded stack per coroutine, each protected from the others.

In ongoing work, we are investigating whether stack protection schemes
based on the Cheri capability system~\citep{Woodruff+14,Chisnall+15}
can enforce our properties.  In particular, we are adapting
the ``uninitialized capability model'' of \citet{Georges+21} to work with
our testing framework in order to validate it against our properties. Out of several Cheri-based
stack protection schemes, this one's treatment of unitialized frames seems most compatible
with our model of confidentiality, without the need for expensive stack clearing
between calls. We expect that Cheri-based techniques can support our properties if we
assume extra cooperation from callers, namely that they do not leak their capabilities
to the attacker. %There is no equivalent dynamic requirement in a tag-based system because
%the policy can dynamically enforce cooperation.
